### 2022-01-01 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_neutral_landscapes_jl@7d7a47946d791dcbc99a0c25675c4acea4a92c36 🚀

Wide range of disciplines: landscape genetics to biogeography.

As biodiversity science becomes increasingly concerned with temporal
change and its consequences, its clear there is a gap generating neutral
landscapes that change over time. In this ms we present how
\texttt{NeutralLandscapes.jl} is orders of magnitudes faster than
packages \texttt{nlmpy} (in python) or \texttt{NLMR} (in R). We then
present a novel method for generating landscape change with prescribed
levels of spatial and temporal autocorrelation, and demonstrate that it
works

@@

fig.~\ref{fig:allmethods} shows a replica of Figure 1 from
(\textbf{nlmpycite?})

@@

\begin{figure}
\hypertarget{fig:allmethods}{%
\centering
\includegraphics{./figures/figure1.png}
\caption{Recreation of the figure in \texttt{nlmpy} paper and the
source}\label{fig:allmethods}
}
\end{figure}

\begin{verbatim}
using NeutralLandscapes, Plots
siz = 50, 50

Fig1a = rand(NoGradient(), siz) # Random NLM
Fig1b = rand(PlanarGradient(), siz) # Planar gradient NLM
Fig1c = rand(EdgeGradient(), siz) # Edge gradient NLM
Fig1d = falses(siz) 
Fig1d[10:25, 10:25] .= true # Mask example
Fig1e = rand(DistanceGradient(findall(vec(Fig1d))), siz) # Mask example
Fig1f = rand(MidpointDisplacement(0.75), siz) # Mask example
Fig1g = rand(RectangularCluster(4, 8), siz)
Fig1h = rand(NearestNeighborElement(200), siz)
Fig1i = rand(NearestNeighborCluster(0.4), siz)
Fig1j = blend([Fig1f, Fig1c])
Fig1k = blend(Fig1h, Fig1e, 1.5)
Fig1l = classify(Fig1i, ones(4))
Fig1m = classify(Fig1a, [1-0.5, 0.5])
Fig1n = classify(Fig1g, [1-0.75, 0.75])
Fig1o = classify(Fig1f, ones(3))
Fig1p = classify(Fig1f, ones(3), Fig1d)
Fig1q = rand(PlanarGradient(90), siz, mask = Fig1n .== 2) #TODO mask as keyword + should mask be matrix or vec or both? (Fig1e)
Fig1r = ifelse.(Fig1o .== 2, Fig1m .+ 2, Fig1o)
Fig1s = rotr90(Fig1l)
Fig1t = Fig1o'

class = cgrad(:Set3_4, 4, categorical = true)
c2, c3, c4 = class[1:2], class[1:3], class[1:4]

gr(color = :fire, ticks = false, framestyle = :box, dpi=500, colorbar = false)
plot(
    heatmap(Fig1a),         heatmap(Fig1b),         heatmap(Fig1c),         heatmap(Fig1d, c = c2), heatmap(Fig1e),
    heatmap(Fig1f),         heatmap(Fig1g),         heatmap(Fig1h),         heatmap(Fig1i),         heatmap(Fig1j), 
    heatmap(Fig1k),         heatmap(Fig1l, c = c4), heatmap(Fig1m, c = c2), heatmap(Fig1n, c = c2), heatmap(Fig1o, c = c3),
    heatmap(Fig1p, c = c4), heatmap(Fig1q),         heatmap(Fig1r, c = c4), heatmap(Fig1s, c = c4), heatmap(Fig1t, c = c3),
    layout = (4,5), size = (1600, 1270)
)

savefig("./figures/figure1.png", )
\end{verbatim}

@@

It's fast. As the scale and resolution of raster data increases, neutral
models must be able to scale to match those data dimensions.

Here we provide two benchmark tests.

First a comparison of the speed variety of methods from each
\texttt{NeutralLandscapes.jl}, \texttt{NLMR}, and \texttt{nlmpy}.

Second we compare these performance of each of these software packages
as rasters become larger. We show that \texttt{Julia} even outperforms
the \texttt{NLMR} via C++ implemention of a particularly slow neutral
landscape method (midpoint displacement).

@@

\textbf{Fig 3} scale comparison

\begin{figure}
\centering
\includegraphics{}
\caption{Comparison of speed of generating a midpoint displacement
neutral landscape (y-axis) against raster size (measured as length of
the size of a square raster, x-axis)}
\end{figure}

>>>>

Wide range of disciplines: from landscape genetics {[}{]}, to spatial
ecology {[}{]}, and biogeography {[}{]}.

As biodiversity science becomes increasingly concerned with temporal
change and its consequences, its clear there is a gap generating neutral
landscapes that change over time. In this ms we present how
\texttt{NeutralLandscapes.jl} is orders of magnitudes faster than
packages \texttt{nlmpy} (in python) or \texttt{NLMR} (in R). In addition
we then present a novel method for generating landscape change with
prescribed levels of spatial and temporal autocorrelation.

@@

fig.~\ref{fig:allmethods} shows a replica of Figure 1 from
(\textbf{nlmpycite?}), which shows the capacity of the library to
generate different types of neutral landscapes, and then apply masks and
categorical classifcation to them.

@@

In

\begin{figure}
\hypertarget{fig:allmethods}{%
\centering
\includegraphics{./figures/figure1.png}
\caption{Recreation of the figure in \texttt{nlmpy} paper and the
source, supplied in less than 40 lines of code.}\label{fig:allmethods}
}
\end{figure}

\hypertarget{what-methods-have-been-called-different-things-but-are-actually-the-same-thing}{%
\subsection{What methods have been called different things but are
actually the same
thing?}\label{what-methods-have-been-called-different-things-but-are-actually-the-same-thing}}

@@

It's fast. As the scale and resolution of raster data increases, neutral
models must be able to scale to match those data dimensions. Here we
provide two benchmark tests. First a comparison of the speed variety of
methods from each \texttt{NeutralLandscapes.jl}, \texttt{NLMR}, and
\texttt{nlmpy}. Second we compare these performance of each of these
software packages as rasters become larger. We show that \texttt{Julia}
even outperforms the \texttt{NLMR} via C++ implemention of a
particularly slow neutral landscape method (midpoint displacement).

@@

In fig 2 we should a selection of neutral landscape generators (random,
edge gradient, perlin noise, distance-gradient)

\hypertarget{mpd-comparison}{%
\subsection{MPD comparison}\label{mpd-comparison}}

Why use this particular generator as the comparison? It's slow. So slow
that NLMR implements it in C++. (NLMR implements both MPD, neighbor,
randrect, and random neighboorhood in c++). Still these three
algorithms, which cosinsts of 3/16 of NLMR's alg implementations,
constitute 33\% of its codebase.

In this section we show our implementation of MPD is faster than NLMR's
C++ MPD across all scales, up to the asyptotic limit imposed by the
\(O(n^2)\) scaling of the raster

\begin{figure}
\centering
\includegraphics{./figures/draftfig3.png}
\caption{Comparison of speed of generating a midpoint displacement
neutral landscape (y-axis) against raster size (measured as length of
the size of a square raster, x-axis)}
\end{figure}
